fastText Supervised Classifier

### 2019-05-02 Code/dallin_annotations/04_build_fasttext_supervised_classifier.Rmd
Made edits and knitted file 04. I used the error = TRUE argument in the code chunks to show code actions previously hidden. Made a new visualization about predictions.

@@

# Load Packages
```{r}
# library(fastrtext)
# library(tidyverse)
```

@@

## List File Paths of Model and Data
To build our fasttext classification model, we begin by defining where we want to store our fastText model once we've trained it. In our case, we want to store the model in Output/models and name the model class_mod_1.dat. Within R, we'll refer to this file path as file_model.

@@

```{r}
# file_model <- "Output/models/class_mod_1.dat"
# train_data_path <- "Output/clean_data/train_data_1.txt"
# test_data_path <- "Output/clean_data/test_data_1.txt"
```

@@

# tweet_data <- read_csv("data/Marketing Research Labeled Tweets_ - tweet_sample_5k_FULL.csv") %>%  
#  filter(nchar(tweet_text) > 5)
```

@@

```{r}
# test_train_data <- 
#   tweet_data %>% 
#   mutate(label = case_when(label == 0 ~ "non-complaint", # each line (tweet) is preceded by a label to be predicted
#                               label == 1 ~ "complaint"),
#             input_lines = paste(label, tweet_text)) %>% 
#   group_by(label) %>% # get a representative sample of both complaints and non complaints
#   mutate(test_train = c("test", "train")[rbinom(n(), 1, 0.8) + 1]) %>% # assign test train splits
#   ungroup()
```

@@

```{r}
# write train data
# train_data_lines <- 
#   test_train_data %>% 
#   filter(test_train == "train") %>% 
#   pull(input_lines) %>% 
#   paste0("__label__", .)
# 
# write_lines(train_data_lines, train_data_path)
# 
# # write test data
# test_data_lines <- 
#   test_train_data %>% 
#   filter(test_train == "test") %>% 
#   pull(input_lines) %>% 
#   paste0("__label__", .)
# 
# write_lines(test_data_lines, test_data_path)

# create a character vector containing the tweets to test without their labels
# test_labels_without_prefix <- 
#   test_train_data %>% 
#   filter(test_train == "test") %>% 
#   pull(label)
```

@@

```{r}
# train model
# execute(commands = c("supervised", "-input", train_data_path, "-output", file_model, "-dim", 20, "-lr", 1, "-epoch", 20, "-wordNgrams", 2, "-verbose", 1))
# 
# # load model
# model <- load_model(file_model)
```

## Review Model:
We'll want to see how a few of the predictions turned out. 
```{r}
# prediction are returned as a list with words and probabilities
# predictions <- predict(model, sentences = test_data_lines, simplify = TRUE, unlock_empty_predictions = TRUE)
# head(predictions, 5)
```

This will help us confirm that our model worked as expected.
```{r}
# length(predictions)
# length(test_labels_without_prefix)
# length(test_data_lines)
```

## Accuracy
Let's look at how accurately the model predicted names
```{r}
# # Compute accuracy
# mean(names(unlist(predictions)) == test_labels_without_prefix, na.rm = TRUE)
```

@@

The hamming loss is the number of wrong labels over the total number of labels. Hamming loss is most useful when applied to multi-classifiers because when it is done with binary classifiers (such as ours), it is simply equal to 1 - accuracy.
```{r}
# # because there is only one category by observation, hamming loss will be the same
# get_hamming_loss(as.list(test_labels_without_prefix), predictions)
# 
# # test predictions
# predictions <- predict(model, sentences = test_to_write)
# print(head(predictions, 5))
# 
# # you can get flat list of results when you are retrieving only one label per observation
# print(head(predict(model, sentences = test_to_write, simplify = TRUE)))
```

>>>>

```{r include=FALSE} knitr::opts_chunk$set(error = TRUE) ```

@@

# Load Packages
```{r}
library(fastrtext)
library(tidyverse)

set.seed(42)
```

@@

## List File Paths of Model and Data
To build our fasttext classification model, we begin by defining where we want to store our fastText model once we've trained it. In our case, we want to store the model in Output/models and name the model class_mod_1.dat. Within R, we'll refer to this file path as model_path.

@@

```{r}
model_path <- here::here("Output", "models", "class_mod_1.dat")
train_data_path <- here::here("Output", "clean_data", "train_data_1.txt")
test_data_path <- here::here("Output", "clean_data", "test_data_1.txt")
```

@@

tweet_data <- read_csv(here::here("Temporary", "Marketing Research Labeled Tweets_ - tweet_sample_5k_Ky-Ch-Ad.csv")) %>%
 filter(nchar(tweet_text) > 5)
```

@@

```{r}
test_train_data <- tweet_data %>%
  mutate(label = case_when(complaint_label == 0 ~ "non-complaint", # each line (tweet) is preceded by a label to be predicted
                              complaint_label == 1 ~ "complaint"),
            input_lines = paste(label, tweet_text)) %>%
  group_by(label) %>% # get a representative sample of both complaints and non complaints
  mutate(test_train = c("test", "train")[rbinom(n(), 1, 0.8) + 1]) %>% # assign test train splits
  ungroup()
```

@@

```{r}
# write train data
train_data_lines <- test_train_data %>%
  filter(test_train == "train") %>%
  pull(input_lines) %>%
  paste0("__label__", .)

write_lines(train_data_lines, train_data_path)

# write test data
test_data_lines <- test_train_data %>%
  filter(test_train == "test") %>%
  pull(input_lines) %>%
  paste0("__label__", .)

write_lines(test_data_lines, test_data_path)

# create a character vector containing the tweets to test without their labels
test_labels_without_prefix <-
  test_train_data %>%
  filter(test_train == "test") %>%
  pull(label)
```

@@

```{r}
# train model
execute(commands = c("supervised", "-input", train_data_path, "-output", model_path, "-dim", 20, "-lr", 1, "-epoch", 20, "-wordNgrams", 2, "-verbose", 1))

# load model
model <- load_model(model_path)
```

## Review Model:
We'll want to see how a few of the predictions turned out. 
```{r}
# prediction are returned as a list with words and probabilities
predictions <- predict(model, sentences = test_data_lines, simplify = TRUE, unlock_empty_predictions = TRUE)

head(predictions, 5)
```

This will help us confirm that our model worked as expected. The predictions are shorter than the test dataset, so I'm not sure what happened. According to some errors I've seen, I think it's because a few tweets use vocabulary that was not included in the training dataset.
```{r}
length(predictions)
length(test_labels_without_prefix)
length(test_data_lines)
```

A second attempt to make predictions:
```{r error = TRUE}
tested <- test_train_data %>% 
  filter(test_train == "test") %>% 
  mutate(prediction = predict(model, tweet_text, unlock_empty_predictions = TRUE)) %>% 
  unnest(prediction) %>% 
  mutate(rounded_prediction = round(prediction),
         correct            = case_when(complaint_label == rounded_prediction ~ TRUE,
                                        complaint_label != rounded_prediction ~ FALSE))
```

Visualization of Outputs:
From this graph, apparently I messed up somewhere in making predictions, because the predictions being made seem to have no correlation with the actual complaint label. It could be that, or the fastText model needs significantly more text to train and be accurate. 
```{r}
tested %>% 
  ggplot(aes(x = row_number(tweet_text), y = prediction, color = complaint_label, alpha = 0.1)) +
  geom_point()
```

Is the list "predictions" the same as the "prediction" variable in tested?
```{r}
are_they_same <- c(tested$prediction[[3]],
                   predictions[[3]],
                   tested$prediction[[300]],
                   predictions[[300]])

(are_they_same)
```
The predictions are exactly the same from both sets, so I guess the classifier was just bad?

I'll try to use our fasttext model using bash.
Navigate to the right directory of the model:
```{bash error = TRUE}
cd ../../Output/models/class_mod_1.bin
ls
```

```{bash error = TRUE}
./fasttext predict ../../Output/models/class_mod_1.bin - @delta you are the worst airline. Thanks for delaying my flight
```
This didn't work

Second attempt to make predictions using bash
```{bash error = TRUE}
cd ../../Output/models

./fasttext predict class_mod_1.bin - @delta you are the worst airline. Thanks for delaying my flight
```

Third attempt to make predictions using bash.
```{bash error = TRUE}
cd ../../../../../../fasttext/fastText-0.2.0
ls
```

f## Accuracy
Let's look at how accurately the model predicted names
```{r error = TRUE}
# Compute accuracy
mean(names(unlist(predictions)) == test_labels_without_prefix, na.rm = TRUE)
```

@@

The hamming loss is the number of wrong labels over the total number of labels. Hamming loss is most useful when applied to multi-classifiers because when it is done with binary classifiers (such as ours), it is simply equal to 1 - accuracy.
```{r error = TRUE}
# because there is only one category by observation, hamming loss will be the same
get_hamming_loss(as.list(test_labels_without_prefix), predictions)

# test predictions
predictions <- predict(model, sentences = test_to_write)
print(head(predictions, 5))

# you can get flat list of results when you are retrieving only one label per observation
print(head(predict(model, sentences = test_to_write, simplify = TRUE)))
```
